---
title: "Mosquito Surveillance Report"
subtitle: "Abundance & West Nile Virus Activity — Weeks 22-26"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    theme: flatly
    highlight: tango
  word_document:
    toc: true
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.height = 5,
  fig.width  = 8
)
```

```{=html}
<style>
  body { font-family: 'Source Sans Pro', 'Segoe UI', sans-serif; font-size: 15px; color: #2c3e50; }
  h1, h2 { color: #1a5276; border-bottom: 2px solid #2e86c1; padding-bottom: 6px; }
  h3 { color: #2471a3; }
  .note-box  { background:#eaf4fb; border-left:4px solid #2e86c1; padding:12px 16px; border-radius:4px; margin:15px 0; }
  .warn-box  { background:#fdedec; border-left:4px solid #e74c3c; padding:12px 16px; border-radius:4px; margin:15px 0; }
  .stat-row  { display:flex; gap:20px; margin:20px 0; flex-wrap:wrap; }
  .stat-card { background:#eaf4fb; border-radius:8px; padding:16px 24px; flex:1; min-width:150px; text-align:center; }
  .stat-card .number { font-size:2em; font-weight:bold; color:#1a5276; }
  .stat-card .label  { font-size:0.9em; color:#566573; margin-top:4px; }
</style>
```

```{r load-packages}
library(vectorsurvR)
library(dplyr)
library(ggplot2)
library(lubridate)
library(kableExtra)
```

```{r config}
# ── USER CONFIGURATION ────────────────────────────────────────────────────────
# Update these values to run the report for your agency and time period

AGENCY_ID  <- 55      # Your agency ID
TARGET_YEAR <- 2020   # Surveillance year
WEEK_START  <- 22     # First epi week of reporting period
WEEK_END    <- 26     # Last epi week of reporting period
IS_HTML     <- TRUE   # Set to FALSE when knitting to Word or PDF

WEEK_SEQ <- as.character(seq(WEEK_START,WEEK_END, by =  1))

# WNV primary vector species for your region — adjust as needed
WNV_SPECIES <- c("Cx tarsalis", "Cx pipiens")

# Derived label used in headings and captions throughout the report
WEEK_LABEL <- paste0("Weeks ", WEEK_START, "\u2013", WEEK_END, ", ", TARGET_YEAR)

# ─────────────────────────────────────────────────────────────────────────────
```

```{r auth, results='hide'}
token <- getToken()
```

```{r load-data, cache=TRUE}
# Collections: pull 5+ years for anomaly comparison
collections_full <- getArthroCollections(
  token,
  start_year = TARGET_YEAR - 5,
  end_year   = TARGET_YEAR,
  arthropod  = "mosquito",
  agency_ids = AGENCY_ID
)

# Current year with epiweek column
collections_year <- collections_full %>%
  filter(year(as.Date(collection_date)) == TARGET_YEAR) %>%
  mutate(epiweek = epiweek(as.Date(collection_date)))

# Filter to reporting week range
collections_period <- collections_year %>%
  filter(epiweek >= WEEK_START & epiweek <= WEEK_END)

# Pools data
pools <- getPools(
  token,
  start_year = TARGET_YEAR,
  end_year   = TARGET_YEAR,
  arthropod  = "mosquito",
  agency_ids = AGENCY_ID
)

# Filter pools to reporting week range
pools_period <- pools %>%
  mutate(epiweek = epiweek(as.Date(collection_date))) %>%
  filter(epiweek >= WEEK_START & epiweek <= WEEK_END)
```

---

## Overview

This report summarizes mosquito surveillance activity and West Nile Virus (WNV) pool testing results for **`r WEEK_LABEL`**. Data reflects collections and pool tests submitted through the VectorSurv Gateway for the reporting agency.

```{r summary-stats}
total_mosquitoes  <- sum(collections_period$num_count, na.rm = TRUE)
total_collections <- n_distinct(collections_period$collection_id)
total_pools       <- n_distinct(pools_period$pool_num)
positive_pools    <- pools_period %>%
  filter(tolower(test_status_name) == "confirmed") %>% nrow()
n_species         <- n_distinct(collections_period$species_display_name)
```

```{=html}
<div class="stat-row">
  <div class="stat-card">
    <div class="number">`r format(total_mosquitoes, big.mark=",")`</div>
    <div class="label">Mosquitoes Collected</div>
  </div>
  <div class="stat-card">
    <div class="number">`r total_collections`</div>
    <div class="label">Collection Events</div>
  </div>
  <div class="stat-card">
    <div class="number">`r n_species`</div>
    <div class="label">Species Detected</div>
  </div>
  <div class="stat-card">
    <div class="number">`r total_pools`</div>
    <div class="label">Pools Tested</div>
  </div>
  <div class="stat-card">
    <div class="number">`r positive_pools`</div>
    <div class="label">WNV Positive Pools</div>
  </div>
</div>
```

```{r wnv-alert, results='asis'}
if (IS_HTML) {
  if (positive_pools > 0) {
    cat('<div class="warn-box">⚠️ <strong>WNV Activity Detected:</strong>',
        positive_pools, 'positive pool(s) confirmed in', WEEK_LABEL,
        '— see the WNV Activity section for details.</div>')
  } else {
    cat('<div class="note-box">✅ <strong>No WNV positives</strong> detected in',
        WEEK_LABEL, 'pool testing.</div>')
  }
} else {
  if (positive_pools > 0) {
    cat("> **⚠️ WNV Activity Detected:**", positive_pools,
        "positive pool(s) confirmed in", WEEK_LABEL)
  } else {
    cat("> **✅ No WNV positives** detected in", WEEK_LABEL)
  }
}
```

---

## Mosquito Abundance

### Abundance by Week — `r WEEK_LABEL`

Abundance is expressed as the average number of mosquitoes caught per trap night.

```{r abundance-table}
getAbundance(
  collections = collections_year,
  interval    = "Week",
  species     = WNV_SPECIES,
  separate_by = "species"
) %>%
  filter(Week >= WEEK_START & Week <= WEEK_END) %>%
  kbl(caption = paste("Table 1: Weekly abundance for primary WNV vector species —", WEEK_LABEL)) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    font_size         = 13,
    latex_options     = "scale_down"
  ) %>%
  row_spec(0, bold = TRUE, background = "#2C3E50", color = "white")
```

---

### Abundance Trend — `r TARGET_YEAR` vs. 5-Year Average

The charts below compare `r TARGET_YEAR` weekly abundance against the historical average for each primary vector species.

```{r abundance-anomaly-plot}
for (sp in WNV_SPECIES) {
  p <- plotVectorAbundance(
    token       = token,
    interval    = "Week",
    target_year = TARGET_YEAR,
    start_year  = TARGET_YEAR - 5,
    end_year    = TARGET_YEAR,
    species     = sp,
    trap        = NULL,
    agency_ids  = AGENCY_ID
  )
  print(p)
}
```

---

### Species Composition — `r WEEK_LABEL`

```{r species-composition}
species_summary <- collections_period %>%
  filter(species_display_name %in% WNV_SPECIES) %>%
  group_by(species_display_name) %>%
  summarise(
    total_count = sum(num_count, na.rm = TRUE),
    .groups     = "drop"
  ) %>%
  arrange(desc(total_count)) %>%
  mutate(pct = round(total_count / sum(total_count) * 100, 1))

ggplot(species_summary, aes(x = reorder(species_display_name, total_count),
                             y = total_count, fill = species_display_name)) +
  geom_bar(stat = "identity", show.legend = FALSE, width = 0.6) +
  geom_text(aes(label = paste0(pct, "%")), hjust = -0.2, size = 3.5, color = "#2c3e50") +
  coord_flip() +
  scale_fill_manual(values = c("#2e86c1", "#1a5276", "#5dade2")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title   = paste("Species Composition —", WEEK_LABEL),
    x       = NULL,
    y       = "Total Mosquitoes Collected",
    caption = "Primary WNV vector species only"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title         = element_text(face = "bold", color = "#1a5276"),
    axis.text.y        = element_text(face = "italic"),
    panel.grid.major.y = element_blank()
  )
```

---

### Collection Map — `r WEEK_LABEL`

```{r species-map}
 plotSpeciesMap(
    token       = token,
    target_year = TARGET_YEAR,
    interval    = "Week",
    time_period = WEEK_SEQ,
    species     = WNV_SPECIES,
    agency_ids  = AGENCY_ID,
    html        = IS_HTML
  )

```

---

## West Nile Virus Activity

### Infection Rate — `r WEEK_LABEL`

Infection rate is estimated using maximum likelihood estimation (MLE), expressed per 1,000 mosquitoes tested.

```{r infection-rate}
getInfectionRate(
  pools,
  interval       = "Week",
  target_disease = "WNV",
  pt_estimate    = "mle",
  scale          = 1000,
  species        = WNV_SPECIES
) %>%
  filter(Week >= WEEK_START & Week <= WEEK_END) %>%
  kbl(caption = paste("Table 2: WNV infection rate —", WEEK_LABEL)) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    font_size         = 13,
    latex_options     = "scale_down"
  ) %>%
  row_spec(0, bold = TRUE, background = "#2C3E50", color = "white")
```

---

### Vector Index — `r WEEK_LABEL`

The vector index combines abundance and infection rate to estimate the relative risk of WNV transmission.

$$\text{Vector Index} = \text{Abundance} \times \text{Infection Rate}$$

```{r vector-index}
getVectorIndex(
  collections_year,
  pools,
  interval       = "Week",
  target_disease = "WNV",
  pt_estimate    = "bc-mle",
  species        = WNV_SPECIES
) %>%
  filter(Week >= WEEK_START & Week <= WEEK_END) %>%
  kbl(caption = paste("Table 3: Vector index —", WEEK_LABEL)) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    font_size         = 13,
    latex_options     = "scale_down"
  ) %>%
  row_spec(0, bold = TRUE, background = "#2C3E50", color = "white")
```

---

### Pool Testing Map — `r WEEK_LABEL`

Positive pools are shown in red; negative pools in blue.

```{r pools-map}
plotPoolsMap(
    token          = token,
    target_year    = TARGET_YEAR,
    interval= "Week",
    target_disease = "WNV",
    time_period    = WEEK_SEQ,
    agency_ids     = AGENCY_ID,
    html           = IS_HTML
  )


```

---

### WNV Positive Pool Summary

```{r positive-pools-table}
if (positive_pools > 0) {
  pools_period %>%
    filter(tolower(test_status_name) == "confirmed") %>%
    select(collection_date, site_code,
           species_display_name, num_count, trap_acronym, epiweek) %>%
    arrange(epiweek, collection_date) %>%
    kbl(
      col.names = c("Date", "Site", "Species", "Mosquitoes", "Trap", "Epi Week"),
      caption   = paste("Table 4: WNV confirmed positive pools —", WEEK_LABEL)
    ) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover"),
      font_size         = 13,
      latex_options     = "scale_down"
    ) %>%
    row_spec(0, bold = TRUE, background = "#c0392b", color = "white")
} else {
  cat("No WNV positive pools were detected in", WEEK_LABEL, ".")
}
```

---

## Summary

```{r written-summary, results='asis'}
wnv_status <- if (positive_pools > 0) {
  paste0("**WNV activity was detected**, with ", positive_pools,
         " confirmed positive pool(s) submitted during this period.")
} else {
  "**No WNV positive pools** were detected during this period."
}

top_species <- species_summary %>% slice_max(total_count, n = 1) %>% pull(species_display_name)

cat(paste0(
  "During **", WEEK_LABEL, "**, surveillance activities recorded **",
  format(total_mosquitoes, big.mark = ","), " mosquitoes** across **", total_collections,
  " collection events**, representing **", n_species, " species**. ",
  "The most abundant WNV vector species was *", top_species, "*. ",
  wnv_status, " A total of **", total_pools, " pools** were submitted for testing. ",
  "Refer to the maps and charts above for spatial and temporal distribution details."
))
```

---

*Report generated using the `vectorsurvR` package. Data sourced from the VectorSurv Gateway.*
